---
title: "Tip of the Month: Correlation Over Time"
output: html_notebook
author: Majeed Simaan
date: May 18, 2018
fig_width: 50
---

In my earlier post from March 2018, I did introduce the `rollapply` function that executes a function on a rolling window basis. While this function is very useful, it needs a little modification for users to apply other general operations.

Originally, I faced this issue when I tried to compute the correlation matrix across different asset returns on a rolling window. For the demonstration, let's consider the returns for all sector ETFs excluding real estate:
```{r message=FALSE, warning=FALSE}
library(quantmod)
v <- c("XLE","XLU","XLK","XLB","XLP","XLY","XLI","XLV","XLF")
t1 <- "1990-01-01"
P.list <- lapply(v,function(x) get(getSymbols(x,from = t1)) )
P.list <- lapply(P.list,function(x) x[,6])
P <- Reduce(merge,P.list)
names(P) <- v
R <- na.omit(P/lag(P) - 1)
```
By default, `rollapply` executes the given function on each time series separately and returns a time series object. For instance, 
```{r message=FALSE, warning=FALSE}
tail(rollapply(R,25,mean))
```
returns the 25 moving average for each one separately. On the other hand, if I try to compute the moving correlation, instead, I get the following
```{r message=FALSE, warning=FALSE}
tail(rollapply(R,25,cor))
```
which computes the correlation with the same ETF rather than other ETFs - as it treats each time series separately. As a remedy, one should add `by.column = F` argument to the `rollapply` function. In this case, the function returns a time series `xts` object, however, with $9 \times 9 = 81$ columns, where each column corresponds to the pairwise correlation between the 9 sector ETFs rather than a squared matrix.
```{r message=FALSE, warning=FALSE}
COR <- rollapply(R,25,cor,by.column = F)
dim(COR)
class(COR)
```
What left to be done is to stack these vectors back into a correlation matrix, one for each time period. To do so, I will refer to the `plyr` package. The `plyr` package allows users to take an array (`a`), a data frame (`d`), or a list (`l`), execute a given function over the given object, and output the results in either format. For our case, I will input the time series `COR` object as an array and output it as a list, where each element in the list  corresponds to the moving correlation matrix. 
```{r message=FALSE, warning=FALSE}
library(plyr)
COR.list <- alply(COR,1,function(x) matrix(x,nrow = ncol(R),  byrow = T ))
```
The second argument in the `alply` specifies the margin, where 1 indicates that the given function to be executed over the rows, while 2 states that it should be executed over the columns instead. The third argument, which takes a function, stacks each row of the `COR` object into a squared matrix. As a result, we have:
```{r message=FALSE, warning=FALSE}
round(COR.list[[25]],2)
```
which is identical to correlation matrix computed over the first 25 days in the data
```{r message=FALSE, warning=FALSE}
round(cor(R[1:25,]),2)
```
Finally, one can either keep the rolling correlation matrix in a list or transform it back a time series using certain computation, e.g. construct portfolio weights and compute the out-of-sample return as a time series. As a finall demostration, I will show how one can stack the list into a time series of average correlation across sectors over time.
```{r message=FALSE, warning=FALSE}
# the following computes average of the upper traingle correlation matrix elements
COR.mean <- sapply(COR.list, function(x) mean(x[upper.tri(x)])  )
summary(COR.mean)
```
To retrieve back into a time series object, following trick should serve well:
```{r message=FALSE, warning=FALSE}
library(lubridate)
names(COR.mean) <- date(COR)
COR.mean <- as.xts(COR.mean) 
plot(COR.mean)
```
Note that, in order to transform a numerical vector into a time series, I label the values with the corresponding date and, then, set it object as an `xts` object, whereas the `lubridate` is an extremely useful package to handle date formats. 




